---
title: "Data 608 Module 3"
author: "Md Jalal Uddin"
date: "March 9, 2019"
output:
  pdf_document: default
  html_document: default
runtime: shiny
---

```{r}
library(shiny)
library(DT)
library(plotly)
library(ggplot2)
library(dplyr)
library(googleVis)
library(rsconnect)

```

Question 1
As a researcher, you frequently compare mortality rates from particular causes across different States. 
You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). 
Create a visualization that allows you to rank States by crude mortality for each cause of death.

```{r}
# Load data
df <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA608/master/lecture3/data/cleaned-cdc-mortality-1999-2010-2.csv", header = TRUE, stringsAsFactors = FALSE)

# Subset for 2010 only
df2010 <- subset(df, Year==2010)

# Unique causes of death for 2010
allcauses <- unique(df2010$ICD.Chapter)

# Create server logic
ui1 <- fluidPage(
  titlePanel("2010 CDC Mortality Rates"),
  fluidRow(selectInput("cause", "Cause of Death:", choices=sort(allcauses))), 
  plotOutput("plot1", height = 800)
)

```


```{r}

# Unique causes of death for 2010
allcauses <- as.data.frame(unique(df2010$ICD.Chapter))

# Create user-interface definition
server1 <- function(input, output) {
  showrates <- reactive({showrates <- subset(df2010, ICD.Chapter==input$cause)})
  output$plot1 <- renderPlot({
    ggplot(showrates(), aes(x=Crude.Rate, y=reorder(State, -Crude.Rate)))+
      scale_x_continuous(limits=c(0, max(showrates()$Crude.Rate))+5, expand = c(0,0))+
      geom_segment(aes(yend=State), xend=0, color="blue")+
      geom_point(size=3, color = "red") +
      theme_bw()+
      theme(panel.grid.major.y = element_blank(), axis.title=element_text(size=14))+
      xlab("2010 CDC Mortality Rate") +
      ylab("State") +
      ggtitle(input$cause)
  }) 
}

```


```{r}
shinyApp(ui = ui1, server = server1)

```
Question 2
Often you are asked whether particular States are improving their mortality rates (per cause) faster than, or slower than, the national average. Create a visualization that lets your clients see this for themselves for one cause of death at the time. Keep in mind that the national average should be weighted by the national population.


```{r}
# Load data
df3 <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA608/master/lecture3/data/cleaned-cdc-mortality-1999-2010-2.csv", header = TRUE, stringsAsFactors = FALSE)

# Get unique lists of inputs
allcauses<-unique(df3$ICD.Chapter)
allstates<-unique(df3$State)

# Create UI script
ui2 <- fluidPage(
  title = "State Mortality Rates Over Time",
  fluidRow(
    column(6, selectInput('causes', 'Cause of Death', choices=sort(allcauses)) ),
    column(6, selectInput('states', 'State', choices=sort(allstates)) )
  ),
  fluidRow(
    plotOutput('myplot')
  )
)

```

```{r}
# Create server script
server2 <- function(input, output) {
  output$myplot <- renderPlot({
    data <- df3 %>% 
      filter(State==input$states, ICD.Chapter==input$causes)
    usavg <- df3 %>% 
      filter(ICD.Chapter==input$causes) %>% 
      group_by(Year) %>% 
      summarise(rateyr=(sum(as.numeric(Deaths))/sum(as.numeric(Population))*100000))
    ggplot(data, aes(x=Year, y=Crude.Rate, color='purple')) + 
      geom_line(size=3) + 
      geom_line(aes(x=usavg$Year, y=usavg$rateyr, color='green'),size=2) + 
      scale_color_manual(
        name='Legend', 
        values=c('green'='green', 'purple'='purple'), 
        labels=c('National Average', 'State'))
  })
}

```

```{r}

shinyApp(ui = ui2, server = server2)

```

